Think deep for reef resilience

### 2025-08-10 paper/paper_draft.qmd
Add quantification of consistently clustered reefs
Also justify why different clustering scales are used/displayed in results section.
Also explain that CID is used to quantify the difference between low and high clusters at management area scale under each GCM for comparisons.

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points). Colour-coded areas outline GBR marine park management regions,
where green represents the Far Northern management region, yellow represents
Cairns/Cooktown, pink covers Townsville/Whitsunday and blue covers Mackay/Capricorn
[@GBRMPA2024]. Reef centroids are marked individually for visual clarity, and are coloured
according to its bioregion [@GBRMPA2001]. All 2429 reefs included in this investigation are
represented in this figure. An annotation represents the approximate location of the
Australian Institute of Marine Science Townsville site at Cape
Cleveland.](../outputs/ADRIA_results/HighResCoralStress/figs/region_map.png){#fig-region-map
fig-align="center" width="14.8cm"}

@@

**\[TI: Ben, best to use the same colours consistently for each management region. Blue,
yellow/orange, green and pink were used in for the map.\]**

@@

Timeseries clustering analyses were performed at the reefal bioregion scale (defined in
GBRMPA [-@GBRMPA2001]). Additional analyses were also performed at larger spatial scales,
such as the GBRMPA management regions and whole-of-GBR scale. These were found to lead to
identical conclusions and so are not shown here. Results of these analyses can be found at
\[**insert citation for repository**\] for those interested.

@@

Under each GCM, in all management areas, low, medium and high cluster reefs have low levels
of distinction between DHW trajectories (@fig-gcm-dhw-trajectories). The highest peak DHW
levels occur under GFDL-CM4 between 2045 and 2050, while greater increases in overall DHWs
timeseries occur under ACCESS-CM2 [@fig-methods-dhws].

![Reef Degree Heating Week timeseries clustered within each management area, for each GCM
analysed. ‘Low’ clusters (green) are clusters with the lowest median area-relative cover
between 2030 and 2040 within each bioregion. ‘Medium’ clusters (orange) are clusters with
intermediate median area-relative cover between 2030 and 2040 within each bioregion. ‘High’
clusters (blue) are clusters with the highest median area-relative cover between 2030 and
2040 within each bioregion. Solid lines represent the cluster median timeseries and bands
represent the 95th percentile confidence intervals. Each management area is represented five
times (for each GCM), with each GCM displayed
horizontally.](../outputs/ADRIA_results/HighResCoralStress/figs/GCM_dhw_timeseries_plot.png){#fig-gcm-dhw-trajectories
fig-align="center" width="14.82cm"}

@@

Comparing across GCMs, it is common for reefs to switch cluster level, including from low to
high and high to low cluster transitions (@sfig-gcm-bioregion-cluster-assignment). This is
expected as spatial differences in GCM DHW timeseries result in reefs being impacted at
different relative magnitudes within their management area across GCMs.

@@

Under all GCMs most reefs are projected to suffer substantial declines in coral cover
relative to 2025 levels. NorESM2-MM and ACCESS-ESM1-5 GCMs convey the most coral cover
maintenance into the future. Differences between low, medium and high reef clusters in cover
timeseries are most notable under these two GCMs [@tbl-cluster-cid;
@fig-gcm-cover-trajectories]. Maintainance of coral cover into the future is linked to the
Equilibrium Climate Sensitivity of the GCM used as levels of future recovery and timespan of
cover maintainance increase as GCM ECS increases [@fig-gcm-cover-trajectories]. Under
high-ECS GCMs (GFDL-CM4, EC-Earth3-Veg and ACCESS-CM2) large declines and lower rates of
recovery in high cover clusters are common compared to lower-ECS GCMs.

@@

>>>>

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points). Colour-coded areas outline GBR marine park management regions,
where orange represents the Far Northern management region, green represents
Cairns/Cooktown, red covers Townsville/Whitsunday and blue covers Mackay/Capricorn
[@GBRMPA2024]. Reef centroids are marked individually for visual clarity, and are coloured
according to its bioregion [@GBRMPA2001]. All 2429 reefs included in this investigation are
represented in this figure. An annotation represents the approximate location of the
Australian Institute of Marine Science Townsville site at Cape
Cleveland.](../outputs/ADRIA_results/HighResCoralStress/figs/region_map.png){#fig-region-map
fig-align="center" width="14.8cm"}

@@

@@

Timeseries clustering analyses were performed at the management area and reefal bioregion
scale (defined in GBRMPA [-@GBRMPA2001]). Additional analyses were also performed at the
whole-of-GBR scale. These were found to lead to identical conclusions and so are not shown
here. To compare timeseries clustering results across GCMs, management area scale clustering
results are used to preserve coarse scale latitudinal differences between regions, while
maintaining ease of comparison and visualisation. These comparisons include quantification
of differences between low and high cover clusters across management regions under each GCM
using the Complexity-Invariant Distance between cluster median timeseries. To assess how
different reef characteristics influence timeseries clustering at a finer scale, bioregion
results are used to capture patterns occurring within individual clusters of reefs. Results
of all analyses at all clustering scales can be found at \[**insert citation for
repository**\] for those interested.

@@

Under each GCM, in all management areas, low, medium and high cluster reefs have low levels
of distinction between DHW trajectories (@sfig-gcm-dhw-trajectories). The highest peak DHW
levels occur under GFDL-CM4 between 2045 and 2050, while greater increases in overall DHWs
timeseries occur under ACCESS-CM2 [@fig-methods-dhws].

@@

Comparing across GCMs, it is common for reefs to switch cluster level, including from low to
high and high to low cluster transitions (@sfig-gcm-bioregion-cluster-assignment). Of 2429
reefs, 20% (503) of reefs have consistent management area scale cluster assignment across
all GCMs. Of the 503 consistently assigned reefs, 49% are within the Far Northern management
area, with 7, 13 and 31% occurring in the Cairns/Cooktown, Townsville/Whitsunday and
Mackay/Capricorn management areas respectively [@sfig-gcm-cluster-map]. Of the 503
consistently assigned reefs, 32, 7 and 61% are clustered into low, medium and high clusters
respectively. High assignment variability across GCMs is expected as spatial differences in
GCM DHW timeseries result in reefs being impacted at different relative magnitudes within
their management area across GCMs.

@@

Under all GCMs many reefs are projected to suffer substantial declines in coral cover
relative to 2025 levels, with an average of 38% of reefs (924 reefs) clustered into the
lowest cover level cluster at the management area scale across GCMs. NorESM2-MM and
ACCESS-ESM1-5 GCMs convey the most coral cover maintenance into the future. Differences
between low, medium and high reef clusters in cover timeseries are most notable under these
two GCMs [@tbl-cluster-cid; @fig-gcm-cover-trajectories]. Maintainance of coral cover into
the future is linked to the Equilibrium Climate Sensitivity of the GCM used as levels of
future recovery and timespan of cover maintainance increase as GCM ECS increases
[@fig-gcm-cover-trajectories]. Under high-ECS GCMs (GFDL-CM4, EC-Earth3-Veg and ACCESS-CM2)
large declines and lower rates of recovery in high cover clusters are common compared to
lower-ECS GCMs.

@@

![Map of the Great Barrier Reef region, including GBRMPA management areas (polygons) and all
considered reefs (points), coloured by management area scale cluster assignment.
Colour-coded areas outline GBR marine park management regions, where orange represents the
Far Northern management region, green represents Cairns/Cooktown, red covers
Townsville/Whitsunday and blue covers Mackay/Capricorn [@GBRMPA2024]. Reef centroids are
marked individually for visual clarity. Reefs that do not have a consistent cluster
assignment level across all GCMs at mangement area scale are gray. All consistently
clustered reefs are coloured according to their management area scale cluster assignment.
Green represents reefs in the low cover cluster, orange represents reefs in the medium cover
cluster and blue represents reefs in the high cover cluster. All 2429 reefs included in this
investigation are represented in this figure. An annotation represents the approximate
location of the Australian Institute of Marine Science Townsville site at Cape
Cleveland.](../outputs/ADRIA_results/HighResCoralStress/figs/gcm_cluster_assignment_map.png){#sfig-gcm-cluster-map
fig-align="center" width="14.82cm"}
